---
title: "Prediction Assignment Writeup"
author: "Leigh Pearson"
date: "2023-11-25"
output:
  html_document: default
  md_document:
    variant: markdown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

For this project, I developed a predictive model to determine the manner in which exercises were performed using data from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants. My approach involved comprehensive data cleaning and preprocessing to handle missing values and ensure data integrity. I employed a Random Forest algorithm, leveraging its robustness in handling large datasets with multiple predictors. The model demonstrated high accuracy and was fine-tuned using cross-validation techniques, ensuring reliability in its predictive capabilities. The final model was tested against a separate dataset of 20 cases, predicting the exercise technique being used. The entire analysis, along with the model and its predictions, have been thoroughly documented and are available in a GitHub repository, showcasing both the methodology and the results in a comprehensible and accessible manner.

## Downloading and Cleaning the Data 
options(repos = c(CRAN = "https://cran.rstudio.com/"))

install.packages("dplyr")

install.packages("tidyr")

install.packages("ggplot2")

```{r, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)

training_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testing_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training_data <- read.csv(training_url)
testing_data <- read.csv(testing_url)
```

Initial exploration reveals many columns with 19216 out of 19622 values missing, which is a very high percentage, and typically, such columns are not useful for predictive modeling or statistical analysis because they contain very little information. Therefore, I will remove these columns from both the training and testing datasets.

```{r}
threshold <- 0.5  # 50% missing values
training_data <- training_data[, colMeans(is.na(training_data)) < threshold]
```
