---
title: "Practical MLS"
author: "Santosh Yadav"
date: "2023-11-25"
output:
  html_document: default
  md_document:
    variant: markdown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The steps are done on 
https://www.coursera.org/learn/practical-machine-learning/lab-sandbox?path=%2F#
and following files were downloaded and put in /home/rstudio 
testing_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
```


They were read via
```
train_data <- read.csv(training_url)
test_data <- read.csv(testing_url)
```

There were 19622 columns and few of them were not proper so stared with 50% threshold. 

```{r}
threshold <- 0.5  # 50% missing values
train_data <- train_data[, columnMeans(is.na(train_data)) < threshold]
```

The step involves were
Step 1 – we need to load the package module 
```
library(gdm)
```
Step 2 – Load the data and if required set a Subset the data of file pml-training.csv
```
data(pml-training.csv)
adData = data.frame(diagnosis,predictors)
```
Step 3 -  fit model
```
fit <- lda(Class~., data= adData[ inTrain,])
```

Step 4 -  summarize the fit
```
summary(fit)
```
Step 5 -  make predictions
```
predictions <- predict(fit, adData)
```
Step 6 -  make predictions
```
predictions <- predict(fit, adData)
```
step 7  summarize accuracy
```
rmse <- mean((adData)^2)
print(rmse)
```

